On the presentation of orthogonal groups over polynomial rings

### 2019-05-13 dl-paper.tex
Rewritten Proposition 4

\begin{rem}
 For a simply-laced $\Phi$ the second assertion of~\cref{thm:Stepanov} can be deduced from the first one by means of~\cref{Zrels}. Indeed, for a root subset $U\subseteq \Phi$ consider the following operation:
 \[dU := U \cup (U - U)\cap \Phi.\]
 In other words $d$ adjoins to $U$ all differences of roots from $U$ which are themselves roots. It is not hard to show that for any parabolic subset $S \subseteq \Phi$ the subset $\Sigma_S$ has the property that $d^n(\Sigma_S) = \Phi$ for some $n$ (in fact, $n=2$). It remains to see that relation~\eqref{Z5} immediately implies the following assertion: every group $G$ containing $\mathcal{Z}(U, R, I)$ also contains $\mathcal{Z}(dU, R, I)$.
\end{rem}

@@

Finally, suppose $g$ has type $5$.
Substituting in~\eqref{Z5} of~\cref{Zrels} $s = -\epsilon f$, $\xi = -\epsilon m$, $\eta=X$ and expressing $z_\alpha(f, mX)$ through other terms we obtain that
\begin{multline} \nonumber z_\alpha(f, mX) = x_{\alpha+\beta}(\epsilon Xf) \cdot x_{\beta}(-mX^2 f) \cdot x_{-\beta}(mf) \cdot x_\alpha(f) \cdot \\ 
 \cdot z_{\alpha+\beta}(-\epsilon X f, -\epsilon m) \cdot z_{-\beta}(-mf, -X) \cdot x_{-\alpha-\beta}(-\epsilon m^2X f) \cdot x_{-\alpha}(-m^2X^2 f), \end{multline}
and the latter expression is a product of generators of type 1, 1, 2, 5, 1, 2, 1, 4.
\end{proof}
For $m \in M$ denote by $P_\alpha(m)$ the coset $P_\alpha(0) \cdot x_{-\alpha}(mX)$.
From the above lemma it follows that $P_\alpha(*)$ coincides with the union of all $P_\alpha(m)$, $m\in M$. This allows us to make the following definition.

@@

\subsection{The subgroups $K(\alpha, \beta)$ and their properties}
\begin{df}
 Let $\alpha, \beta$ be a pair of nonopposite roots of $\Phi$. Denote by $K(\alpha, \beta)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ 
 generated by the following elements:
 \begin{enumerate}
  \item $z_\gamma(Xf, \xi)$, $\gamma \in \Phi \setminus\{ \pm \alpha, \pm \beta\};$
  \item $x_{-\alpha}(X^2f)$, $x_{-\beta}(X^2f)$;
  \item $x_{\alpha}(Xf)$, $x_\beta(Xf)$;
  \item $c_{\pm \delta}(f, X\xi)$ for some fixed root $\delta \in \Phi$ forming a sharp angle with both $\alpha$ and $\beta$.
 \end{enumerate}
\end{df}
\begin{lemma} \label{root-lemma}
 For any $\alpha, \beta$ in a system of type $A_3$ there exist $\alpha_1, \alpha_2, \beta_1, \beta_2 \in \Phi$ such that
 $\alpha = \alpha_1 - \alpha_2$, $\beta = \beta_1 - \beta_2$ and root sets $\{ \pm \alpha, \pm \beta \}$ and $\{ \alpha_1, \alpha_2, \beta_1, \beta_2 \}$ have empty intersection.
\end{lemma}
\begin{proof} Direct calculation. \end{proof}

\begin{prop} \label{K-a-b}
 Let $g$ be an element of $K(A[X], M[X])$, then for any pair of roots $\alpha, \beta \in \Phi$, $\alpha \neq \pm \beta$
 the element $g$ can be written as a product $g_0 \cdot x_{-\alpha}(mX) \cdot x_{-\beta}(m'X)$ for some $m, m' \in M$ and $g_0 \in K(\alpha, \beta)$.
\end{prop}
\begin{proof}
 Embed $\alpha$ and $\beta$ in some subsystem $\Psi$ of type $A_3$ and denote by $\Sigma$ the special part of the parabolic subset $S = \Psi \cup \Phi^+$.
 Applying~\cref{thm:Stepanov} to $S$ we obtain that every element of $\St(\Phi, A[X], XM[X])$ can be presented as a product of 
  $z_\gamma(Xf, \xi)$, $\gamma \in \Sigma(S) \subseteq \Phi\setminus \{ \pm \alpha, \pm \beta \}$ and $x_\gamma(Xf)$, $\gamma \in \{ \pm \alpha, \pm \beta \}$.

 By~\cref{Kgen-strong} we can write $g$ as a product of elements of the above family and generators of the form $c_{\pm \delta}(f, X\xi)$ for $\delta$
  as in the above definition.

 The next step of the proof is to decompose each factor of the form $x_{\epsilon}(Xf)$, $\epsilon \in \{-\alpha,-\beta\}$ appearing in this product as
 $x_{\epsilon}(X^2f') \cdot x_{\epsilon}(mX)$ and then move $x_{\epsilon}(mX)$ to the rightmost position
  (conjugating by $x_{\epsilon}(mX)$ all factors along the way).
 Notice that no terms of the form $x_{\epsilon}(mX)$, $\epsilon \in \{ -\alpha, -\beta \}$ may appear after the simplification of conjugates,
  so this process is guaranteed to terminate.
 Indeed, notice that the extra terms in the right-hand sides of \eqref{eq3-1} and~\eqref{eq3-2} 
 (as well as all terms in the right-hand sides of~\eqref{C1},\eqref{C2} of~\cref{Crels} if one substitutes corresponding values into them)
  have terms of degree at least $2$ in $X$.

 As a side effect of this transformation, however, factors $x_\alpha(Xf)$ and $x_\beta(Xf)$ appearing in $g$ might be transformed into terms of the form
   $z_\alpha(Xf, X\xi)$, and $z_\beta(Xf, X\xi)$. Applying preceding lemma to the subsystem $\Psi$ we can write $\alpha = \alpha_1 - \alpha_2$ and 
    $\beta = \beta_1 - \beta_2$ with the additional property $\{ \pm \alpha, \pm \beta \} \cap \{ \alpha_1, \alpha_2, \beta_1, \beta_2 \} = \emptyset$.
 Invoking~\eqref{Z5} of~\cref{Zrels} with $\alpha = \alpha_1$, $\beta = -\alpha_2$, $\eta = 1$, $s = Xf$, $\xi = X\xi$ we can rewrite the term
  $z_\alpha(Xf, X\xi)$ as a product of generators of type 1, 2, 3. Similar argument also applies to $z_\beta(Xf, X\xi)$.

@@

\begin{lemma}
 Let $\alpha, \beta \in \Phi$ be such that $\langle \alpha, \beta \rangle = 0\text{ or }1$.
 Then the image of $K(\alpha, \beta)$ under $S_{\alpha}(a, -)$ is contained in the subgroup of $\St(\Phi, A[X, X^{-1}])$ 
 generated by $K(\alpha, \beta)$ and the subgroup $U$, by definition, generated by $z_\gamma(f, X\xi)$, $\gamma \in Z_+(\alpha) \setminus \{ \alpha - \beta \}$.
 We call these elements generators of type 5.
\end{lemma}
\begin{proof}
By~\eqref{eq:Zconj1}, \eqref{eq:Zconj2} the conjugate to each generator $z_\gamma(Xf, \xi)$ of type 1 by $x_\alpha(aX^{-1})$ is a product of generators of type 5, 5, 1.
By~\eqref{eq:Zconj3} the conjugate to the generator $x_{-\alpha}(X^2f)$ is a product of generators of type 1, 5, 1, 2, 1, 1, 5, 5.
$S(\alpha, -)$ either fixes $x_{-\beta}(X^2f)$, the other generator of type 2, or transforms them into products of generators of type 2, 1.
Clearly, $S(\alpha, -)$ fixes both generators of type 3. 
As for the remaining generators type 4, \cref{Crels} shows that the corresponding conjugate can be written as a product of generators of type 5, 5, 4.
\end{proof}

>>>>

\begin{rem}
 For a simply-laced $\Phi$ the second assertion of~\cref{thm:Stepanov} can be deduced from the first one by means of~\cref{Zrels}. Indeed, for a root subset $U\subseteq \Phi$ consider the following operation:
 \[dU := U \cup (U - U)\cap \Phi.\]
 In other words, $d$ adjoins to $U$ all differences of roots from $U$ which are themselves roots. It is not hard to show that for any parabolic subset $S \subseteq \Phi$ the subset $\Sigma_S$ has the property that $d^n(\Sigma_S) = \Phi$ for some $n$ (in fact, for $n=2$). It remains to see that relation~\eqref{Z5} immediately implies the following assertion: every group $G$ containing $\mathcal{Z}(U, R, I)$ also contains $\mathcal{Z}(dU, R, I)$.
\end{rem}

@@

Finally, suppose $g$ has type $5$.
Substituting in~\eqref{Z5} of~\cref{Zrels} $s = -\epsilon f$, $\xi = -\epsilon m$, $\eta=X$ and expressing $z_\alpha(f, mX)$ through other terms we obtain that
\begin{multline} \label{eq:zalpha} z_\alpha(f, mX) = x_{\alpha+\beta}(\epsilon Xf) \cdot x_{\beta}(-mX^2 f) \cdot x_{-\beta}(mf) \cdot x_\alpha(f) \cdot \\ 
 \cdot z_{\alpha+\beta}(-\epsilon X f, -\epsilon m) \cdot z_{-\beta}(-mf, -X) \cdot x_{-\alpha-\beta}(-\epsilon m^2X f) \cdot x_{-\alpha}(-m^2X^2 f), \end{multline}
and the latter expression is a product of generators of type 1, 1, 2, 5, 1, 2, 1, 4.
\end{proof}
For $m \in M$ denote by $P_\alpha(m)$ the coset $P_\alpha(0) \cdot x_{-\alpha}(mX)$.
From the above lemma it follows that $P_\alpha(*)$ coincides with the union of all $P_\alpha(m)$, $m\in M$. This allows us to make the following definition.

@@

\subsection{The subgroups $K(\alpha, \beta)$ and their properties}
Throughout this section $\alpha, \beta$ denote a fixed pair of roots of $\Phi$ forming a sharp angle (i.\,e. $\langle \alpha, \beta \rangle = 1$). We denote by $\Psi$ the subsystem of type $\rA_2$ generated by $\alpha$ and $\beta$.
\begin{df}
 Denote by $K(\alpha, \beta)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by the following elements:
 \begin{enumerate}
  \item $z_\gamma(Xf, \xi)$, for all $\gamma \in \Phi \setminus \Psi,$ and $x_{\alpha-\beta}(Xf)$, $x_{\beta-\alpha}(Xf)$;
  \item $x_{-\alpha}(X^2f)$, $x_{-\beta}(X^2f)$;
  \item $x_{\alpha}(Xf)$, $x_\beta(Xf)$;
  \item $c_{\pm \delta}(f, X\xi)$ for some fixed root $\delta \in \Phi \setminus \Psi$.
 \end{enumerate}
\end{df}

\begin{prop} \label{K-a-b}
 Every element $g \in K(A[X], M[X])$ can be written as a product $g_0 \cdot x_{-\alpha}(mX) \cdot x_{-\beta}(m'X)$ for some $m, m' \in M$ and $g_0 \in K(\alpha, \beta)$.
\end{prop}
\begin{proof}
 Denote by $\Sigma$ the special part of the parabolic subset $S = \Psi \cup \Phi^+$.
 Applying~\cref{thm:Stepanov} to $S$ we obtain that every element of $\St(\Phi, A[X], XM[X])$ can be presented as a product of 
  $z_\gamma(Xf, \xi)$, $\gamma \in \Sigma(S) \subseteq \Phi\setminus \Psi$ and $x_\gamma(Xf)$, $\gamma \in \Psi$.
 By~\cref{Kgen-strong} we can write $g$ as a product of these generators and generators of type 4.

 The next step of the proof is to decompose each factor of the form $x_{\epsilon}(Xf)$, $\epsilon \in \{-\alpha,-\beta\}$ appearing in this product as
 $x_{\epsilon}(X^2f') \cdot x_{\epsilon}(mX)$ and then move $x_{\epsilon}(mX)$ to the rightmost position
  (conjugating by $x_{\epsilon}(mX)$ all factors along the way).
 Notice that no terms of the form $x_{\epsilon}(mX)$, $\epsilon \in \{ -\alpha, -\beta \}$ may appear after the simplification of conjugates,
  so this process is guaranteed to terminate.
 To see this notice that the extra terms in the right-hand sides of \eqref{eq3-1} and~\eqref{eq3-2} 
 (as well as all terms in the right-hand sides of~\eqref{C1},\eqref{C2} of~\cref{Crels} if one substitutes corresponding values into them)
  have terms of degree at least $2$ in $X$.

 As a side effect of this transformation, however, factors $x_\alpha(Xf)$ and $x_\beta(Xf)$ appearing in $g$ might transform into factors of the form $z_\alpha(Xf, X\xi)$, and $z_\beta(Xf, X\xi)$. We can express each factor $z_\alpha(Xf, X\xi)$ through the generators of $K(\alpha, \beta)$ as follows. Pick a root $\gamma \in \Phi\setminus \Psi$ forming an obtuse angle with $\alpha$ and then invoke~\eqref{eq:zalpha}:
 \begin{multline}
  z_\alpha(Xf, X\xi) = x_{\alpha+\gamma}(\epsilon X^2f) \cdot x_{\gamma}(-X^3 \xi f) \cdot x_{-\gamma}(Xf\xi) \cdot x_\alpha(Xf) \cdot \\ \cdot z_{\alpha+\gamma}(-\epsilon X^2 f, -\epsilon \xi) \cdot z_{-\gamma}(-Xf\xi, -X) \cdot x_{-\alpha-\gamma}(-\epsilon X^2\xi^2 f) \cdot x_{-\alpha}(-X^3 f \xi^2),\ \text{for } \epsilon = N_{\alpha, \gamma}.
 \end{multline}
 The expression in the right-hand side is a product of generators of type 1, 1, 1, 3, 1, 1, 1, 2. Similar argument also applies to $z_\beta(Xf, X\xi)$.

@@

\begin{lemma}
 The image of $K(\alpha, \beta)$ under $S_{\alpha}(a, -)$ is contained in the subgroup of $\St(\Phi, A[X, X^{-1}])$ 
 generated by $K(\alpha, \beta)$ and the subgroup $U$, which, by definition, is generated by $z_\gamma(f, X\xi)$, $\gamma \in Z_+(\alpha) \setminus \{ \alpha - \beta \}$.
 We call these elements generators of type 5.
\end{lemma}
\begin{proof}
By~\eqref{eq:Zconj1}, \eqref{eq:Zconj2} the conjugate to each generator $z_\gamma(Xf, \xi)$ of type 1 by $x_\alpha(aX^{-1})$ is a product of generators of type 5, 5, 1.
By~\eqref{eq:Zconj3} the conjugate to the generator $x_{-\alpha}(X^2f)$ is a product of generators of type 1, 5, 1, 2, 1, 1, 5, 5.
$S(\alpha, -)$ either fixes $x_{-\beta}(X^2f)$, the other generator of type 2, or transforms them into products of generators of type 2, 1.
Clearly, $S(\alpha, -)$ fixes both generators of type 3. 
As for the remaining generators type 4, \cref{Crels} shows that the corresponding conjugate can be written as a product of generators of type 5, 5, 4.
\end{proof}
